---
title: "MTCars Performance Analysis"
author: "Your Name"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  message = FALSE,
  warning = FALSE,
  fig.width = 7,
  fig.height = 4
)
library(tidyverse)
library(broom)
library(GGally)
```

# Introduction

This report explores the classic built-in **`mtcars`** dataset in R, focusing on:

- Relationships between **fuel efficiency** (`mpg`) and engine characteristics
- Correlations among numerical variables
- A simple **linear regression model** for predicting `mpg`

The dataset contains information on 32 car models from the 1970s.

# 1. Inspecting the data

```{r}
mtcars %>% 
  as_tibble(rownames = "model") %>% 
  head()
```

```{r}
summary(mtcars)
```

# 2. Exploratory Data Analysis

We’ll focus on a few key variables:

- `mpg` – miles per gallon (target)
- `hp` – gross horsepower
- `wt` – weight (1000 lbs)
- `cyl` – number of cylinders

## 2.1 Distribution of mpg

```{r}
mtcars %>% 
  ggplot(aes(x = mpg)) +
  geom_histogram(bins = 10, color = "white") +
  labs(
    title = "Distribution of Fuel Efficiency (mpg)",
    x = "Miles per gallon",
    y = "Count"
  )
```

## 2.2 mpg vs horsepower

```{r}
mtcars %>% 
  ggplot(aes(x = hp, y = mpg)) +
  geom_point() +
  geom_smooth(method = "lm", se = TRUE) +
  labs(
    title = "Fuel efficiency vs horsepower",
    x = "Horsepower",
    y = "Miles per gallon"
  )
```

## 2.3 mpg vs weight

```{r}
mtcars %>% 
  ggplot(aes(x = wt, y = mpg)) +
  geom_point() +
  geom_smooth(method = "lm", se = TRUE) +
  labs(
    title = "Fuel efficiency vs vehicle weight",
    x = "Weight (1000 lbs)",
    y = "Miles per gallon"
  )
```

# 3. Correlation exploration

```{r}
mtcars %>% 
  select(mpg, cyl, disp, hp, wt, qsec) %>% 
  ggpairs(title = "Correlation plot for selected variables")
```

# 4. Linear regression model

We’ll build a simple model to predict `mpg` using `hp` and `wt`:

```{r}
model <- lm(mpg ~ hp + wt, data = mtcars)
summary(model)
```

Tidy summary:

```{r}
tidy(model)
glance(model)
```

**Interpretation (example):**

- The coefficient for `wt` shows how much `mpg` is expected to change for each additional 1000 lbs, holding horsepower constant.
- The coefficient for `hp` shows how `mpg` changes per extra unit of horsepower, holding weight constant.
- The R-squared value tells us how much of the variation in `mpg` is explained by these predictors.

# 5. Model diagnostics

```{r}
par(mfrow = c(2, 2))
plot(model)
par(mfrow = c(1, 1))
```

From these diagnostic plots we can inspect:

- **Residuals vs Fitted**: check for non-linearity or heteroscedasticity
- **Normal Q-Q**: check if residuals are approximately normal
- **Scale-Location** and **Residuals vs Leverage**: identify any influential points

# 6. Conclusion

In this mini-project we:

- Explored how **horsepower** and **weight** relate to fuel efficiency.
- Saw that heavier, more powerful cars tend to have lower `mpg`.
- Built a simple linear regression model and inspected its diagnostics.

This structure mirrors a small, portfolio-ready R analysis and can be easily extended with:

- Additional predictors (e.g. `cyl`, `am`)
- Interaction terms
- Model selection (AIC, stepwise)
